Supplementary material to:  On the propriety of restricted Boltzmann machines

### 2016-12-01 writing/appendix/appendix-note.Rmd
nordman and vardeman final edits

{\bf Proof of Proposition \ref{prop:instab}.} We prove the contrapositive, supposing that $\Delta(\boldsymbol \theta_N) \le C$ holds for some $C > 0$ and show $\text{ELPR}(\boldsymbol \theta_N) \leq C$. Let $\boldsymbol x_{min} \equiv \argmin\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$ and $\boldsymbol x_{max} \equiv \argmax\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$. Note there exists a sequence $\boldsymbol x_{min} \equiv \boldsymbol x_0, \boldsymbol x_1, \dots, \boldsymbol x_k \equiv \boldsymbol x_{max}$ in $\mathcal{X}^N$ of component-wise switches to move from $\boldsymbol x_{min}$ to $\boldsymbol x_{max}$ in the sample space (i.e. $\boldsymbol x_i, \boldsymbol x_{i + 1} \in \mathcal{X}^N$ differ in exactly $1$ component for $i = 0, \dots, k$) for some integer $k \in \{0, 1, \dots, N\}$. Under the FSFS model, recall $P_{\boldsymbol \theta_N}(\boldsymbol x) > 0$ holds so that $\log P_{\boldsymbol \theta_N}(\boldsymbol x)$ is well-defined for each outcome $\boldsymbol x \in \mathcal{X}^N$. Then, if $k > 0$, it follows that
$$
N\cdot\text{ELPR}(\boldsymbol \theta_N) = \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] = \left|\sum\limits_{i = 1}^k\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta_N}(\boldsymbol x_{i-1})}\right)\right| \le \sum\limits_{i = 1}^k\left|\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta}(\boldsymbol x_{i-1})}\right)\right| \le k \Delta_N(\boldsymbol \theta_N) \le NC,
$$
using $k \le N$ and $\Delta(\boldsymbol \theta_N) \le C$. If $k = 0$, then $\boldsymbol x_{max} = \boldsymbol x_{min}$ and the same bound above holds.$\Box$

>>>>

{\bf Proof of Proposition \ref{prop:instab}.} We prove the contrapositive, supposing that $\Delta(\boldsymbol \theta_N) \le C$ holds for some $C > 0$ and show $\text{ELPR}(\boldsymbol \theta_N) \leq NC$. Let $\boldsymbol x_{min} \equiv \argmin\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$ and $\boldsymbol x_{max} \equiv \argmax\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)$. Note there exists a sequence $\boldsymbol x_{min} \equiv \boldsymbol x_0, \boldsymbol x_1, \dots, \boldsymbol x_k \equiv \boldsymbol x_{max}$ in $\mathcal{X}^N$ of component-wise switches to move from $\boldsymbol x_{min}$ to $\boldsymbol x_{max}$ in the sample space (i.e. $\boldsymbol x_i, \boldsymbol x_{i + 1} \in \mathcal{X}^N$ differ in exactly $1$ component for $i = 0, \dots, k$) for some integer $k \in \{0, 1, \dots, N\}$. Under the FSFS model, recall $P_{\boldsymbol \theta_N}(\boldsymbol x) > 0$ holds so that $\log P_{\boldsymbol \theta_N}(\boldsymbol x)$ is well-defined for each outcome $\boldsymbol x \in \mathcal{X}^N$. Then, if $k > 0$, it follows that
$$
\text{ELPR}(\boldsymbol \theta_N) = \log\left[\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_{max})}{P_{\boldsymbol \theta_N}(\boldsymbol x_{min})}\right] = \left|\sum\limits_{i = 1}^k\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta_N}(\boldsymbol x_{i-1})}\right)\right| \le \sum\limits_{i = 1}^k\left|\log\left(\frac{P_{\boldsymbol \theta_N}(\boldsymbol x_i)}{P_{\boldsymbol \theta}(\boldsymbol x_{i-1})}\right)\right| \le k \Delta_N(\boldsymbol \theta_N) \le NC,
$$
using $k \le N$ and $\Delta(\boldsymbol \theta_N) \le C$. If $k = 0$, then $\boldsymbol x_{max} = \boldsymbol x_{min}$ and the same bound above holds.$\Box$
